Molecular dynamics simulation of synchronization in driven particles

### 2021-02-15 manuscript/manuscript_ajp_2020.tex
updated abstract and intro text.  filled in phase diagram results - not analysis.  likewise on hops

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  Modeling 
  single particle
  synchronization 
  with simulations or table-top experiments
  can is a useful technique to isolate
  complex behaviors %to 
  %provide insight
  in the natural world.
  We discuss a %simple
  %numerical model %that
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as we alter the landscape and driving force.
  Under certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked.
  We include molecular dynamics code %and exercises
  to simulate and characterize
  particle dynamics.
  %including 
  %plotting  particle velocity as a function of applied driving force 
  %and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of synchronized 
pendula in wall-mounted clocks \cite{Bennett2002}.
%Even when started out of phase, 
%Huygens demonstrated the tendency of
%coupled pendula tend to swing in time
%due to interactions through the wall 
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks,
%even when started out of phase, 
the pendula were observed swing at same rate -- a 1:1 mode.
For simple pendula of different lengths,
a 2:1 mode can be seen when a pendulum four
times the length of another swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena

@@

Complex dynamics can be studied with
colloid particles in experiments or simulations.
Typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
%experimental - colloids specifically
Because 
colloids are large %-- a few microns in size
and move slowly %-- a few microns per second,
particle position 
can be captured in real time 
with an optical camera \cite{Pertsinidis2001}. %(see Table ~\ref{tab:1}).
Experimental measurements of 
step-by-step dynamics of 
colloids performing phase-locked motions
are useful for
understanding synchronization at a single particle level \cite{Juniper2015}.

%due to the large size and slow speed of the particles.
%Individual colloids exhibit 
%Brownian motion,
%where particles change direction %in a seemingly random fashion
%due to 
%collisions 
%with unseen particles comprising the suspending fluid \cite{}.
%[elaborate - move to associated problems?].

Light is a
tool for manipulating the colloidal environment.
A circular optical trap uses 
radiation pressure from 
a laser beam to confine colloids \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in the trap is %subject to a
uniformly bombarded by photons. % while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
Off-center colloids %away from center
experience a net force
due to uneven photon collisions across
the particle surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward center 
%local minimum 
or ejects it from the trap.
%In experiments
Diffraction gratings can be used to create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat,
 or moving between substrate minima.
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 A constant or 
 dc drive the landscape modulates 
 the particle velocity.
 Below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 A sinusoidal or ac drive can create mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.
 When the applied force varies periodically,
 the ac drive can 
 cause the particle to hop back and forth across
 the landscape minima.
 Many synchronized patterns occur
 controlled by the substrate period.

@@

The molecular dynamics simulation is controlled by a $for()$ loop
which runs from an initial to maximum integer number of time steps.
%$t_i$.
Each integer time step interval %$t_{i+1}-t_i$
represents a simulation time of $\Delta t$ 
simulation units $\tau$.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position.
%$\vec{r}_i(t+\Delta t)$.%
Since the acceleration as zero,
the integration of the equation of motion
is performed via 
the Euler method 
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)
    \label{eq:euler}
\end{equation}
for a small time step $\Delta t = 0.001 \tau$.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%
The units of the simulated variables are summarized in Table ~\ref{tab:1}.

@@

A single particle %on this landscape
responds to the applied driving force
by moving across the landscape %maxima
synchronized with 
the period of $F^D(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = N_p \pi U_0/L$.
In this section, we fix the landscape parameters
to $A_{p} = 0.1$ force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$
to illustrate
distinct oscillation modes. % as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
%as shown

  It is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot.
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When $v_y$ %a particle velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  $v_y$ vs $y$ space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a $1-1$ mode appears as a circle or oval.
  Nodes appear in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is not fully phase locked
  will apppear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized.
  %and the shifts in phase space represent
  %small changes in system behavior.
  With a long simulation time,
  a quasiperiodic phase diagram
  fills the plotting region due to the gradual shift
  in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

  In a system driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\bar{v}_y t)/\lambda
  \end{equation}
  centered about the average particle velocity $\bar{v}_y$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  We plot this phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\bar{v}_y/\lambda.
  \end{equation}
  %Juniper %REF 53

  In Fig.~\ref{fig:phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing substrate amplitude $A_p$.
  We fix the driving force with parameters
  $F^{dc} = 0.1$ and frequency $f = 0.01$ and $F^{ac} = 0.2$.
  In Fig.~\ref{fig:phase}(a)
  there is no substrate ($A_p = 0.0$)
  and the phase plot is a closed oval
  centered at
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$ %ANALYTIC
  and semiminor axis of $\Delta \dot{\phi} = 2\pi/\lambda v_{max}$.
  [insert analysis.]
  As the force from the substrate
  increases,
  nodes form in the phase diagram
  and the motion becomes quasiperiodic
  so the loops in phase space do not fully overlap.
  In Fig.~\ref{fig:phase}(b)
  $A_p = 0.05$
  the spatial properties are similar to
  those in $A_p = 0.0$
  with 
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$.
  The dynamical properties
  are asymmetric,
  with nodes in the positive $\dot{\phi}$ values.
  [insert analysis.]
  As the substrate
  force becomes larger,
  $\phi$ shifts to lower values
  and $\dot{\phi}$ to positive values,
  indicating
  the particle can only move 
  backward over a short distance.
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{phase.pdf}
      \caption{
        Phase plot of a particle with the same values of $F^{dc}$ and frequency as Fig.~\ref{fig:0}.  $F^{ac} = 0.2$ and $A_p = $ (a) $0.0$, (b) $0.05$, (c) $0.1$, (d) $0.2$, (e) $0.25$ and (f) $0.3$.
      }
      \label{fig:phase}
    \end{figure}
  \end{center}

In Fig.~\ref{fig:1} 
we keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$,
with $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
%We plot the results in Fig.~\ref{fig:1} for 
The modes appear as 
discrete steps in $\langle v_y \rangle$ %occur because
across a range of $F^{dc}$,

@@

We explore a range of parameters in Ex.~\ref{ex:parameters}
where 
we perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve. 
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as non-linear 
steps in the force-velocity relationship.

@@

  {\bf With no applied driving force,
    find 
  the minimum temperature required for a single particle
  to hop over maxima in the potential landscape.}
  %$T_{min} > A_p$.
  Assume the particle is confined to
  move along the $y-$direction,
  include Brownian motion 
  in the model.
  In Fig.~\ref{fig:brownian},
  we demonstrate the transition
  from pinned to hopping
  in 
  %Measure the position vs. time
  a non-driven colloid for $T/A_p = 4.0, 5.0, 6.0$.
  The crossover
  to a hopping regime occurs at $T/A_p \approx 5.6$.
  %Reproduce these results.  
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{{brownian_particle_dt0.1}.pdf}
      \caption{
        Motion of a particle undergoing Brownian motion
        on a washboard potential
        energy landscape.
        (a) $T/A_p = 4.0$ no hopping occurs.
        (b) $T/A_p = 5.0$ a three hops occur %at $t = 130000$
        across distances of $\Delta y = 2 \lambda$.
        (c) $T/A_p = 6.0$ many hops occur.
        Note the scale of the $y-$axis differs in each panel
        and the time scale is
        much greater than in Fig.~\ref{fig:0}.
      }
      \label{fig:brownian}
    \end{figure}
  \end{center}
  In Fig.~\ref{fig:brownian}
  we show the position versus time of a particle
  confined to a sinusoidal landscape
  undergoing Brownian motion.
  We increase the temperature relative
  to the amplitude of the landscape
  troughs until the particle enters a hopping regime.
  In  Fig.~\ref{fig:brownian}(a)
  the temperature is $T/A_p = 4.0$.
  The particle executes a 
  random walk centered at the potential minima $y/\lambda = 10$.
  We ran many simulations and never observed a hop to another minima
  within this simulation time.  
  In  Fig.~\ref{fig:brownian}(b)
  $T/A_p = 5.0$ where hops between
  potential minima are possible but not probable,
  a single hop occurs from
  $y/\lambda = 10$ to $y/\lambda = 12$ at
  $t \sim 40000 \tau$ and 
  $y/\lambda = 10$ to $y/\lambda = 8$ at
  $t \sim 280000 \tau$.
  In other simulations with identical parameters,
  we sometimes observed no hops or two hops between minima,
  as expected with random fluctuating systems.
  In  Fig.~\ref{fig:brownian}(c)
  $T/A_p = 6.0$ 
  we observe many hops.
  The random thermalized kicks are
  sufficiently large to make the particle perform
  what appears to be 
  a random walk of hops atop the substrate.
  %in addition to the small scale random walk executed
  %by all particles in this section.

@@

  %(b)
  For a driven particle,
  we ignore 
  the effects of temperature 
  in these simulations.  
  We note that even at $T/A_p = 6.0$ 
  the hopping rate
  is much less than the
  frequency of the applied drive.
  This can be seen in the time scale in Fig.~\ref{fig:brownian},
  where the average hopping rate is approximately
  every $15000\tau$,
  a value much larger than $T = 100\tau$.
  At sufficiently high temperatures,
  Brownian motion does affect 
  the formation of Shapiro steps
  and can be observed in experiments.
  %and could be included
  %to emulate noise seen in some 
  %How does Brownian motion affect
  %With driving force reproduce the paper figures...
  %%[there is no way I want to run this with python...]

>>>>

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  %Modeling
  A particle may
  execute
  motion 
  coordinated  with position,
  a synchronization behavior
  that 
  can be used to isolate
  complex natural behaviors,  %to
  either with
  simulations or table-top experiments.
  %provide insight
  We discuss a %simple
  %numerical model %that
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as we alter the landscape and driving force.
  Under certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked,
  forming closed orbits in phase space.
  Quasi-periodic motion is common, % in this system,
  where the
  dynamical center of motion shifts the
  phase space orbit.
  We include molecular dynamics code %and exercises
  to simulate and characterize
  particle dynamics.
  %including 
  %plotting  particle velocity as a function of applied driving force 
  %and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of synchronized 
pendula in wall-mounted clocks \cite{Bennett2002}.
%Even when started out of phase, 
%Huygens demonstrated the tendency of
%coupled pendula tend to swing in time
%due to interactions through the wall 
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks,
%even when started out of phase, 
the pendula were observed swing at same rate -- a 1:1 mode.
For a higher modes,
consider simple pendula of different lengths,
a 2:1 mode occurs when a pendulum four
times the length of another swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena

@@

Complex dynamics such as synchronized mode-locking
can be studied with
colloid particles in experiments or simulations.
Typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
%experimental - colloids specifically
Because 
colloids are large %-- a few microns in size
and move slowly, %-- a few microns per second,
particle position 
can be measured in real time 
with an optical camera \cite{Pertsinidis2001}. %(see Table ~\ref{tab:1}).
Experimental measurements of 
step-by-step dynamics of 
colloids performing phase-locked motions
are useful for
understanding synchronization at a single particle level \cite{Juniper2015}. 
%due to the large size and slow speed of the particles.
%Individual colloids exhibit 
%Brownian motion,
%where particles change direction %in a seemingly random fashion
%due to 
%collisions 
%with unseen particles comprising the suspending fluid \cite{}.
%[elaborate - move to associated problems?].

Light is a
tool for manipulating the colloidal environment
to alter synchronization patterns.
Colloids can be trapped
with 
%A circular optical trap uses 
radiation pressure from 
a laser beam \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in an optical trap is %subject to a
uniformly bombarded by photons. % while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
Off-center colloids %away from center
experience a net force
due to uneven photon collisions across
the particle surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward center 
%local minimum 
or ejects it from the trap.
%In experiments
Diffraction gratings can create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat,
 or moving between substrate minima.
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 A constant or 
 dc drive the landscape modulates 
 the particle velocity.
 Below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 When the applied force varies periodically,
 the ac drive can 
 cause the particle to hop back and forth across
 the landscape minima.
 Many synchronized patterns occur
 controlled by the substrate period,
 leading to 
 %A sinusoidal or ac drive can create
 mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.

@@

The molecular dynamics simulation is controlled by a $for()$ loop
which runs from an initial to maximum integer number of time steps.
%$t_i$.
Each integer time step interval %$t_{i+1}-t_i$
represents a simulation time of $\Delta t$ 
simulation units $\tau$.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position.
%$\vec{r}_i(t+\Delta t)$.%
Since the acceleration as zero,
the integration of the equation of motion
is performed via 
the Euler method 
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)
    \label{eq:euler}
\end{equation}
for a small time step $\Delta t = 0.01 \tau$.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%
The units of the simulated variables are summarized in Table ~\ref{tab:1}.

@@

A single particle %on this landscape
responds to the applied driving force
by moving across the landscape %maxima
synchronized with 
the period of $F^D(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = N_p \pi U_0/L$.
In this section, we fix the landscape parameters
to $A_{p} = 0.1 F_0$ %force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
  It is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot.
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  In a system driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\langle v_y \rangle t)/\lambda
  \end{equation}
  centered about the average particle velocity $\langle v_y \rangle$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  We plot this phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\langle v_y \rangle /\lambda.
  \end{equation}
  %Juniper %REF 53
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  phase space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a 1:1 mode appears as a circle or oval.
  Nodes appear in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is not fully phase locked
  will apppear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized.
  %and the shifts in phase space represent
  %small changes in system behavior.
  With a long simulation time,
  a quasiperiodic phase diagram
  fills the plotting region due to the gradual shift
  in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

  In Fig.~\ref{fig:phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing substrate amplitude $A_p$.
  We fix the driving force with parameters
  $F^{dc} = 0.1$ and frequency $f = 0.01$ and $F^{ac} = 0.2$.
  In Fig.~\ref{fig:phase}(a)
  there is no substrate ($A_p = 0.0$)
  and the phase plot is a closed oval
  centered at
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$ %ANALYTIC
  and semiminor axis of $\Delta \dot{\phi} = 2\pi/\lambda v_{max}$.
  [insert analysis.]
  As the force from the substrate
  increases,
  nodes form in the phase diagram
  and the motion becomes quasiperiodic
  so the loops in phase space do not fully overlap.
  In Fig.~\ref{fig:phase}(b)
  $A_p = 0.05$
  the spatial properties are similar to
  those in $A_p = 0.0$
  with 
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$.
  The dynamical properties
  are asymmetric,
  with three nodes in the positive $\dot{\phi}$ values.
  The thickening of the line
  indicates the system is not fully synchronized.
  [insert analysis.]
  As the substrate
  force becomes larger,
  $\phi$ shifts to lower values
  and $\dot{\phi}$ to positive values,
  indicating
  the particle can only move 
  backward over a short distance.
  In Fig.~\ref{fig:phase}(c)
  $A_p = 0.1$
  the phase properties are similar to
  those in $A_p = 0.05$,
  the three nodes have 
  increased asymmetry in both phase variables 
  $(\phi,\dot{\phi})$.
  The nodes deepen,
  indicating [insert analysis.].
  In Fig.~\ref{fig:phase}(d)
  $A_p = 0.2$
  the pattern shifts
  to two nodes,
  [indicating change analysis].
  In Fig.~\ref{fig:phase}(e)
  $A_p = 0.25$
  the system returns to a nodeless
  orbit,
  shaped like a triangle [because why?]
  In Fig.~\ref{fig:phase}(f)
  $A_p = 0.3$
  the particle is trapped between substrate minima,
  and traces out a small unclosed
  orbit in phase space [because...]
  The phase velocity is near zero
  throughout the motion -
  [comment on initial conditions]
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{phase.pdf}
      \caption{
        Phase plot of a particle with the same values of $F^{dc}$ and frequency as Fig.~\ref{fig:0}.  $F^{ac} = 0.2$ and $A_p = $ (a) $0.0$, (b) $0.05$, (c) $0.1$, (d) $0.2$, (e) $0.25$ and (f) $0.3$.
      }
      \label{fig:phase}
    \end{figure}
  \end{center}

To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$
to illustrate
distinct oscillation modes. % as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
%as shown
In Fig.~\ref{fig:1} 
we keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$,
with $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
%We plot the results in Fig.~\ref{fig:1} for 
The modes appear as 
discrete steps in $\langle v_y \rangle$ %occur because
across a range of $F^{dc}$,

@@

We explore a range of parameters in Ex.~\ref{ex:parameters}
where 
we perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve. 
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
%We examine
The period of the substrate
can be used to control the
net velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as non-linear 
steps in the force-velocity relationship.

@@

  {\bf With no applied driving force,
    find 
  the minimum temperature required for a single particle
  to hop over maxima in the potential landscape.}
  %$T_{min} > A_p$.
  Assume the particle is confined to
  move along the $y-$direction,
  include Brownian motion 
  in the model.
  In Fig.~\ref{fig:brownian},
  we demonstrate the transition
  from pinned to hopping
  in 
  %Measure the position vs. time
  a non-driven colloid for $T/A_p = 4.0, 5.0, 6.0$.
  The crossover
  to a hopping regime occurs at $T/A_p \approx 5.6$.
  %Reproduce these results.  
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{{brownian_particle_dt0.1}.pdf}
      \caption{
        Motion of a particle undergoing Brownian motion
        on a washboard potential
        energy landscape.
        (a) $T/A_p = 4.0$ no hopping occurs.
        (b) $T/A_p = 5.0$ a three hops occur %at $t = 130000$
        across distances of $\Delta y = 2 \lambda$.
        (c) $T/A_p = 6.0$ many hops occur.
        Note the scale of the $y-$axis differs in each panel
        and the time scale is
        much greater than in Fig.~\ref{fig:0}.
      }
      \label{fig:brownian}
    \end{figure}
  \end{center}
  In Fig.~\ref{fig:brownian}
  we show the position versus time of a particle
  confined to a sinusoidal landscape
  undergoing Brownian motion.
  We increase the temperature relative
  to the amplitude of the landscape
  troughs until the particle enters a hopping regime.
  In  Fig.~\ref{fig:brownian}(a)
  the temperature is $T/A_p = 4.0$.
  The particle executes a 
  random walk centered at the potential minima $y/\lambda = 10$.
  We ran many simulations and never observed a hop to another minima
  within this simulation time.  
  In  Fig.~\ref{fig:brownian}(b)
  $T/A_p = 5.0$,  hops between
  potential minima are possible but not probable,
  a back-forth hop occurs from
  $y/\lambda = 10$ to $12$ at
  $t \sim 40000 \tau$ and a single hop from 
  $y/\lambda = 10$ to $8$ occurs at
  $t \sim 280000 \tau$.
  In other simulations with identical parameters,
  we sometimes observed no hops or two hops between minima,
  as expected with random fluctuating systems.
  In  Fig.~\ref{fig:brownian}(c)
  $T/A_p = 6.0$ 
  we observe many hops.
  The random thermalized kicks are
  sufficiently large to make the particle perform
  what appears to be 
  a random walk of hops atop the substrate.
  [prove it]
  %in addition to the small scale random walk executed
  %by all particles in this section.

@@

  %(b)
  For a driven particle,
  we ignore 
  the effects of temperature 
  in these simulations.  
  We note that even at $T/A_p = 6.0$ 
  the hopping rate
  is much less than the
  frequency of the applied drive.
  This can be seen in the time scale in Fig.~\ref{fig:brownian},
  where the average hopping rate is approximately
  every $15000\tau$,
  a value much larger than $T = 100\tau$.
  At sufficiently high temperatures,
  Brownian motion does affect 
  the formation of mode-locked steps
  and can be observed in experiments.
  [explore]
  %and could be included
  %to emulate noise seen in some 
  %How does Brownian motion affect
  %With driving force reproduce the paper figures...
  %%[there is no way I want to run this with python...]
